\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-01-29 forwards-and-backwards.tex
compact containment

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

Now we prove the main theorem on convergence of the lookdown process, 
Theorem~\ref{thm:lookdown_convergence}.
This follows a similar pattern as the the proofs of convergence for the population process
in Section~\ref{sec:population_density_proof},
although we will need additional estimates to ensure that
trajectories of individual lines of descent converge,
not just the population process.

\begin{proof}[Proof of Theorem~\ref{thm:lookdown_convergence}]
    As in Section~\ref{sec:population_density_proof},
    the theorem will follow from tightness,
    ensuring convergence along subsequences,
    and characterization of the limit.
    This time, the processes $\lp^N$ take values in $\lpmeasures$,
    the space of locally finite measures on space $\times$ levels.
    (They will in fact be point measures, including the limit,
    but that is a consequence of this theorem.)
    Again, tightness will follow from a compact containment condition,
    tightness of one-dimensional distributions,
    and an application of \citet{EK} Theorem~3.9.1.
    We show that the martingale properties are preserved in the limit
    in Section~\ref{sec:lookdown_generator_proofs}.
    \comment{TODO: write and refer to lemma?}

    Lines of descent can escape to infinite level in finite time,
    and so we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    i.e., the topology induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $f(x) g(u)$, where $f \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $g \in C_c([0,\infty))$ is continuous with compact support.
    In this topology, compact containment follows from
    tightness of the spatial and level motions of a single line of descent
    (Lemmas~\ref{lem: tightness of individual spatial trajectory}
    and~\ref{lem:tightness of levels}, respectively),
    and
    tightness of $\lp^N(\IR^d \times [0, u_0])$ for each $u_0$
    (i.e., the number of individuals with level below $u_0$).

    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    However, by Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    This implies that any limit of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.

\comment{MAKE THIS BELOW A LEMMA?}

% BEGIN OLD STUFF
To prove that the lookdown representation is tight, we generalise the vague
    topology to the setting of measure-valued processes. For a detailed
    description of the vague topology, refer to Appendix \ref{sec: Empirical
    Measure Topology}.

We will lay out the steps we follow to prove tightness. First of all, we will
    show that compact containment condition is satisfied if we can control the
    total mass of particles in our system living in a finite region in space
    with a level lower than some $u_{*}$. Since a control on total mass has
    already been shown in Lemma \ref{lem:eta_compact_containment}, we have
    immediately established compact containment condition for our lookdown
    processes.

Then, we apply Theorem A 2.4 in \cite{kallenberg1997foundations} (Theorem
    \ref{teo: Vague Relative Compactness through Projection}) and use a
    corollary of Stone-Weierstrass Theorem (Lemma \ref{lem: Dense set for test
    functions on empirical measures}) to show that our lookdown representation
    $\{\xi^N_t: N \in \mathbb{N}\}$ is tight as a $\mathcal{M}(\IR^d
    \times [0, \infty])$-valued stochastic process in the vague topology as
    long as $\{F_g(\xi^N_t): N \in \mathbb{N}\}$ is tight as real-valued
    stochastic processes for a huge class of $F_g$. Explicitly, $F_g(\xi):=
    \prod_{(x_i,u_i)\in \xi}g(x_i,u_i)$, where $g(x,u)$ satisfies the property
    that $g(x,u) = 1 $ for all $u \geq u_{*}$ for some $u_{*}>0$.
% END OLD STUFF

    The process we consider is similar to the state-dependent branching processes
    of \citet{KR}, so one might expect that the proofs there would carry over with little change.
    However, there is an important difference:
    Recall that the level $U_a(t)$ of a line of descent evolves as
    $$
    \dot u 
    =
    c(x, \eta)
    u^2
    - b(x, \eta)
    u,
    $$
    where
    \begin{align*}
        c(x, \eta) &= \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) \\
        \text{and}\qquad
        b(x, \eta)
        &=
        \theta \gamma(x, \eta) \int_{\IR^d} \left(r(y, \eta) - r(x, \eta)\right) q(x, dy)
        + F(x, \eta) 
    \end{align*}
    (see equations \eqref{eqn:level_generator} and \eqref{eqn:b_limit}),
    where $c(x, \eta) \ge 0$ and $b(x, \eta)$ may take either sign.
    In \citet{kurtz/rodrigues:2011}, $b$ was bounded above and $c$ was bounded away from zero,
    so they noted that if $U_a(t) \ge b/c$
    lines of descent would only move upwards from that time onwards.
    This allows the processes to be jointly and simultaneously constructed
    for all values of $N$, with a pointwise embedding
    of $(\xi^N_t)_{t \ge 0}$ within $(\xi^{M})_{t \ge 0}$ for $b/c < N < M$.
    In other words, individuals with levels above $N > b/c$ at time $t_0$
    do not affect $(\lp^N_t)_{t \ge t_0}$,
    thus allowing a comparison of the number of lines of descent below level $u_0$
    to a branching process.
    Although we have provided a
    joint construction of $\lp^N$ for all $N$ in Section~\ref{sec: individual lines of descent},
    it does not have this monotonicity:
    for one thing, $b$ and $c$ depend on the population process $\eta$
    and so all individuals can affect all other ones (even those with lower levels).
    Furthermore, in the deterministic case $\theta/N$ and hence $c$ converge to zero,
    and so lines of descent with arbitrarily high level may drift back downwards.
    Indeed, this must be the case if the population persists,
    since in the deterministic case there is no branching.
    In our case $b$ and $c$ depend on the state of the population,
    so we have no such monotonicity.

@@

% % % % % % % % % % % % %
\subsubsection{New stuff}

\comment{TODO: reference this elsewhere, if we keep it?}
We record for future reference
the Aldous-Rebolledo criteria for tightness of a sequence real-valued processes:

\begin{theorem}[\citet{rebolledo:1980}]
    \label{thm:aldous_rebolledo}
    Let $\{Y^(n)\}_{n \ge 1}$ be a sequence of real-valued processes
    with c\`adl\`ag paths.
    Suppose that the following conditions are satisfied.
    \begin{enumerate}
        \item For each fixed $t$, $\{Y_t^(n)\}_{n \ge 1}$  is tight.
        \item Given a sequence of stopping times $\tau_n$, bounded by $T$,
            for each $\epsilon > 0$ there exists $\delta > 0$ and $n_0$ such that
            \begin{align*}
                \sup_{n \ge n_0}
                \sup_{\theta \in [0, \delta]}
                \IP\left\{ \left|
                    Y^{(n)}_{\tau_n + \theta} - Y^{(n)}_{\tau_n}
                \right| > \epsilon
                \right\}
                \le \epsilon .
            \end{align*}
    \end{enumerate}
    Then the sequence $\{Y^(n)\}_{n \ge 1}$ is tight.
\end{theorem}

>>>>

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{C}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR^d})} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on space x level

@@

Now we turn to the main theorem on convergence of the lookdown process, 
Theorem~\ref{thm:lookdown_convergence}, whose proof
follows a similar pattern as the the proofs of convergence for the population process
in Section~\ref{sec:population_density_proof}.

The process we consider is similar to the state-dependent branching processes
of \citet{KR}, so one might expect that the proofs there would carry over with little change.
However, there is an important difference:
Recall that the level $U_a(t)$ of a line of descent evolves as
$$
\dot u 
=
c(x, \eta)
u^2
- b(x, \eta)
u,
$$
where
\begin{align*}
    c(x, \eta) &= \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) \\
    \text{and}\qquad
    b(x, \eta)
    &=
    \theta \gamma(x, \eta) \int_{\IR^d} \left(r(y, \eta) - r(x, \eta)\right) q(x, dy)
    + F(x, \eta) 
\end{align*}
(see equations \eqref{eqn:level_generator} and \eqref{eqn:b_limit}),
where $c(x, \eta) \ge 0$ and $b(x, \eta)$ may take either sign.
In \citet{kurtz/rodrigues:2011}, $b$ was bounded above and $c$ was bounded away from zero,
so they noted that if $U_a(t) \ge b/c$
lines of descent would only move upwards from that time onwards.
This allows the processes to be jointly and simultaneously constructed
for all values of $N$, with a pointwise embedding
of $(\lp^N_t)_{t \ge 0}$ within $(\lp^{M})_{t \ge 0}$ for $b/c < N < M$.
In other words, individuals with levels above $N > b/c$ at time $t_0$
do not affect $(\lp^N_t)_{t \ge t_0}$,
thus allowing a comparison of the number of lines of descent below level $u_0$
to a branching process.
Although we have provided a
joint construction of $\lp^N$ for all $N$ in Section~\ref{sec: individual lines of descent},
it does not have this monotonicity:
for one thing, $b$ and $c$ depend on the population process $\eta$
and so all individuals can affect all other ones (even those with lower levels).
Furthermore, in the deterministic case $\theta/N$ and hence $c$ converge to zero,
and so lines of descent with arbitrarily high level may drift back downwards.
Indeed, this must be the case if the population persists,
since in the deterministic case there is no branching.
In our case $b$ and $c$ depend on the state of the population,
so we have no such monotonicity.

\begin{proof}[Proof of Theorem~\ref{thm:lookdown_convergence}]
    As in Section~\ref{sec:population_density_proof},
    the theorem will follow from tightness and characterization of the limit.
    This time, the processes $\lp^N$ take values in $\lpmeasures$,
    the space of locally finite measures on space $\times$ levels.
    (They will in fact be point measures, including the limit,
    but that is a consequence of this theorem.)

    Again, tightness follows from a compact containment condition
    (Lemma~\ref{lem:compact containment lookdown}),
    tightness of one-dimensional distributions,
    and an application of \citet{EK} Theorem~3.9.1.
    Lines of descent can escape to infinite level in finite time,
    and so we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    i.e., the topology induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0,\infty))$ is continuous with compact support.
    Theorem~3.9.1 of \citet{EK} requires that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes for all $F$ in a subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, so
    by the Stone-Weierstrass theorem, the algebra they generate is dense in $C_0(\lpmeasures)$,
    the continuous functions that vanish at infinity
    (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    Since $C_0(\lpmeasures)$ is dense in $\bar C(\lpmeasures)$
    with respect to uniform convergence on compact subsets,
    it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.

    We show that the martingale properties are preserved in the limit
    in Section~\ref{sec:lookdown_generator_proofs}.
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    However, by Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    This implies that any limit of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

Let $C_c^+$ denote the set of nonnegative continuous functions on $\overline{\IR^d} \times [0, \infty)$
for which there exists a $u_f \ge 0$ such that $u > u_f$ implies that $f(x, u) = 0$.
Recall that we are working with the vague topology ``in the level direction'',
so that $m_n \to m$ in $\lpmeasures$ if and only if $\langle f, m_n \rangle \to \langle f, m \rangle$
for all $f \in C_c^+$.
By Theorem~A.23 in \citet{kallenberg1997foundations},
the space $\lpmeasures$ is separable and metrizable,
and $K \subset \lpmeasures$ is relatively compact if and only if
$\sup_{\lp \in K} \langle f, \lp \rangle < \infty$
for any $f \in C_c^+$.
% The compact sets we consider are those with bounded total mass of their spatial projection,
% i.e., of the form
% $$
%     K_M = \left\{
%         \lp \in \lpmeasures \;:\;
%         \sup_{L \ge 1} \frac{1}{L} \lp\left(\overline{\IR^d} \times [0, L]\right) \le M
%     \right\} .
% $$
% These sets are compact 
% because if $\lp \in K_M$ and $f \in C_c^+$ then $\langle f, \lp\rangle \le M u_f \|f\|_\infty$.
Let $\mathcal{H} = \{h_n\}_{n \ge 1}$ be a sequence of
monotone, nonnegative, compactly supported, differentiable functions
on $[0, \infty)$ such that there exist sequences $u_n \to \infty$ and $c_n$ such that
\begin{enumerate}
    \item $h_n(u) > 1$ for $u \le u_n/2$,
    \item $h_n(u) = 0$ for $u \ge u_n$,
    \item $- h_n(u) \le h_n'(u) \le 0$, and
    \item $\max\left( \int_u^\infty h_n(v) dv, \int_u^\infty h_n(v)^2 dv \right) \le c_n h_n(u)$.
\end{enumerate}
For instance, we could take $h_n(u)$ to be a mollified version of $\min(1, e^{-(u-n)})$,
and then $u_n = c_n = n$.
The compact sets we consider are those with uniformly bounded integrals against $\mathcal{H}$,
i.e., of the form
$$
    K_C = \left\{
        \lp \in \lpmeasures \;:\;
        \langle h_n, \lp \rangle \le C_n
        \text{ for all } n \ge 1
    \right\} ,
$$
for some sequence of constants $(C_1, C_2, \ldots)$.
This set is compact because any $f \in C_c^+$ is bounded by a constant multiple of some $h_n$.
\comment{check: why is it closed?}

\begin{lemma}[Compact containment for $\lp$]
    \label{lem:compact containment lookdown}
    Under Assumptions~\ref{},
    if $\theta/N \to \alpha$ and $\lp_0^N \to \lp_0$ weakly as $N \to \infty$
    \comment{and they are conditionally Poisson},
    for any $T$ and $\delta > 0$ there exists a sequence
    $(C_1, C_2, \ldots)$ of finite constants such that
    \begin{align} \label{eqn:compact eqn lookdown}
    \limsup_{N \to \infty}
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
        \text{ for some } n \ge 1
    \right\}
    < \delta .
    \end{align}
\end{lemma}

In other words, the processes $\lp^N$ stay in $K_C$ for all $0 \le t \le T$
with uniformly high probability.
We will prove this Lemma using the following one:

\begin{lemma}
    \label{lem:compact containment hn}
    Under the assumptions of Lemma~\ref{lem:compact containment lookdown},
    for any $T$, $M$ and $n$ there exists a constant $K_n$ such that
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > M \right\}
    <
    \frac{K_n}{M} .
    $$
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:compact containment lookdown}]
    By a union bound,
    $$ 
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
        \text{ for some } n \ge 1
    \right\}
    \le
    \sum_{n \ge 1}
    \IP\left\{
        \sup_{0 \le t \le T} \langle h_n, \lp^N_t \rangle > C_n
    \right\} ,
    $$
    so \eqref{eqn:compact eqn lookdown} follows
    by taking $C_n = 2^{n-1} K_n / \delta$ in Lemma~\ref{lem:compact containment hn}.
\end{proof}

\begin{proof}[of Lemma~\ref{lem:compact containment hn}]
    \comment{TODO: just give bounds on moments and say this follows
        the proof of Lemma~\ref{lem:eta_compact_containment} exactly, otherwise.
        Or make the basic idea a lemma in the appendix?}
    Let $h \in \mathcal{H}$.
    It is evident from the description of the process
    (or, by differentiating in Definition~\ref{defn:lookdown_mgale})
    that
    \begin{align*}
    \begin{split}
        \langle h^n(u), \lp_t^N(dx, du) \rangle
        &=
        \langle h^n(u), \lp_0^N(dx, du) \rangle
        + M^{h_n}_t 
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            2 a^N(x, \eta^N_s) \int_u^N h_n(u_1) du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h_n'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
    \end{split}
    \end{align*}
    where $M^{h_n}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h_n}]_t
        &=
        \int_0^t
        \langle 2 a^N(x, \eta_s^N) \int_u^N h(u_1)^2 du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le a^N(x, \eta_x^N) \le C_a$ for some $C_a < \infty$,
    and we have assumed that $\int_u^\infty h_n(u_1) du_1 \le C_n h_n(u)$,
    so that 
    \begin{align*}
        & \int_0^t
        \big\langle
            2 a^N(x, \eta^N_s) \int_u^N h(u_1) du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad
        \le
        2 C_a C_n \int_0^t \big\langle h_n(u), \lp_s^N(dx, du) \big \rangle .
    \end{align*}
    In the next line, the contribution from $a^N(x, \eta^N_s) u^2 h'(u)$ is nonpositive because
    $a^N \ge 0$ and $h'(u) \le 0$.
    Similarly, since $b^N(x, \eta_x^N) \le C_b$ for some $C_b < \infty$,
    and by assumption
    $- h'(u) \le C_n h(u)$ and $h'(u) = 0$ for $u \ge u_n$,
    and so
    \begin{align*}
        & \int_0^t
        \big\langle
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h_n'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \le
        C_b u_n C_n \int_0^t \big\langle h_n(u), \lp_s^N(dx, du) \big \rangle .
    \end{align*}
    Putting this together, we have that
    \begin{align} \label{eqn:hn_bound}
        \langle h^n, \lp_t^N \rangle
        &\le
        \langle h^n, \lp_0^N \rangle
        + M^{h_n}_t 
        + (2 C_a + C_b u_n) C_n \int_0^t 
            \langle h^n, \lp_s^N \rangle
        ds .
    \end{align}
    By Gronwall's inequality, with $C = (2 C_a + C_b u_n) C_n$
    \begin{align*}
        \IE\left[ \langle h^n, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h^n, \lp_0^N \rangle \right] e^{Ct} .
    \end{align*}

    Now consider the maximum. By \eqref{eqn:hn_bound},
    \begin{align*}
        \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle
        &\le
        \langle h^n, \lp_0^N \rangle
        + \sup_{0 \le t \le T} M^{h_n}_t 
        + C \int_0^t 
            \sup_{0 \le t \le T} \langle h^n, \lp_s^N \rangle
        ds .
    \end{align*}
    Again since $\sqrt{x} \le 1 + x$ for $x \ge 0$,
    The Burkholder-Davis-Gundy inequality
    tells us that there is a $C'$ such that
    \begin{align*}
    \IE\left[ \sup_{0 \le t \le T} M^{h_n}_t \right]
    & \le
        C'\left( 1 + \IE\left[ [M^{h_n}]_t \right] \right)
    \\ & \le
        C' \left( 1 + \int_0^t \IE\left[
        \langle 2 a^N(x, \eta_s^N) \int_u^N h(u_1)^2 du_1
        ,
        \lp_s^N(dx, du) \big \rangle
        \right] ds \right) 
    \\ & \le
        C' \left( 1 + 2 C_a C_n \int_0^t \IE\left[
        \langle h_n, \lp_s^N \big \rangle
        \right] ds \right) 
    \\ & \le 
        C' \left( 1 + C'' e^{Ct} \IE[ \langle h^n, \lp_t^N \rangle ] \right) ,
    \end{align*}
    where $C'' = 2 C_a C_n / C$.
    Therefore, 
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle \right]
        &\le
        C' + \IE\left[ \langle h^n, \lp_0^N \rangle \right]( 1 + C'' e^{Ct} )
        + C \int_0^t 
            \IE\left[ \sup_{0 \le t \le T} \langle h^n, \lp_s^N \rangle \right]
        ds ,
    \end{align*}
    and so once again by Gronwall's inequality, there is a $C'''$ such that
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle \right]
        &\le
        C'''(1 + \IE[\langle h^n, \lp_0^N \rangle]) e^{2Ct} ,
    \end{align*}
    and so
    \begin{align*}
        \IP\left\{ \sup_{0 \le t \le T} \langle h^n, \lp_t^N \rangle > K \right\}
        &\le
        \frac{
            C'''(1 + \IE[\langle h^n, \lp_0^N \rangle]) e^{2CT} ,
        }{ K } .
    \end{align*}
\end{proof}

We record for future reference
the Aldous-Rebolledo criteria for tightness of a sequence real-valued processes:
\comment{TODO: reference this elsewhere, if we keep it?}

\begin{theorem}[\citet{rebolledo:1980}]
    \label{thm:aldous_rebolledo}
    Let $\{Y^(n)\}_{n \ge 1}$ be a sequence of real-valued processes
    with c\`adl\`ag paths.
    Suppose that the following conditions are satisfied.
    \begin{enumerate}
        \item For each fixed $t$, $\{Y_t^{(n)}\}_{n \ge 1}$  is tight.
        \item Given a sequence of stopping times $\tau_n$, bounded by $T$,
            for each $\epsilon > 0$ there exists $\delta > 0$ and $n_0$ such that
            \begin{align*}
                \sup_{n \ge n_0}
                \sup_{\theta \in [0, \delta]}
                \IP\left\{ \left|
                    Y^{(n)}_{\tau_n + \theta} - Y^{(n)}_{\tau_n}
                \right| > \epsilon
                \right\}
                \le \epsilon .
            \end{align*}
    \end{enumerate}
    Then the sequence $\{Y^{(n)}\}_{n \ge 1}$ is tight.
\end{theorem}

\begin{lemma}
    \label{lem:product_tightness}
    Let $\{X^{(n)}\}_{n \ge 1}$ and $\{Y^{(n)}\}_{n \ge 1}$ sequences of jointly defined
    real-valued processes with c\`adl\`ag paths
    satisfying the conditions of Theorem~\ref{thm:aldous_rebolledo}.
    Then $\{X^{(n)}Y^{(n)}\}_{n \ge 1}$ and $\{X^{(n)} + Y^{(n)}\}_{n \ge 1}$ also
    satisfy the conditions of Theorem~\ref{thm:aldous_rebolledo}.
\end{lemma}

By ``jointly defined'' we mean that $X^{(n)}$ and $Y^{(n)}$ are defined on the same
probability space, so that the products and sums make sense.

\begin{proof}[of Lemma~\ref{lem:product_tightness}]
    The proof for $X^{(n)} + Y^{(n)}$ is similar to but more straightforward than
    for $X^{(n)} Y^{(n)}$, so on only prove the Lemma for the latter.
    First, note that for any $\epsilon > 0$, by tightness of $X^{(n)}_t$ and $Y^{(n)}_t$
    there is a $K$ such that
    $\IP\{ X^{(n)}_t > \sqrt{K} \}$ and $\IP\{ Y^{(n)}_t > \sqrt{K} \}$
    are both less than $\epsilon/2$,
    and hence
    \begin{align*}
        \IP\{
            X^{(n)}_t Y^{(n)}_t > K
        \}
        \le
        \IP\{
            X^{(n)}_t > \sqrt{K}
        \}
        +
        \IP\{
            Y^{(n)}_t > \sqrt{K}
        \} 
        \le \epsilon.
    \end{align*}
    Therefore, $X^{(n)}_t Y^{(n)}_t$ is tight.

    Next, note that for $0 \le \tau_n \le T$,
    \begin{align*}
        &
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        \\ &\qquad 
        \le
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left| X^{(n)}_{\tau_n + \theta} \right|
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \left| Y^{(n)}_{\tau_n} \right|
        \\ &\qquad 
        \le
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \sup_{0 \le t \le T}
        \left| Y^{(n)}_t \right| ,
    \end{align*}
    so that for any $C$,
    \begin{align} \label{eqn:xy_bound}
        \begin{split}
        &
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        > \epsilon
        \right\}
    \\ &\qquad \le
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| X^{(n)}_t \right|
            > C
        \right\}
    +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    \\ &\qquad {} +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    +
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| Y^{(n)}_t \right|
            > C
        \right\}
        \end{split}
    \end{align}
    Now, since $\max_{0 \le t \le T} X^{(n)}_t$ is tight (and likewise for $Y$)
    (see, e.g., Remark 3.7.3 in \citet{EK}),
    we may choose a $C \ge 4$ for which
    $$
    \IP\left\{
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        > C
    \right\} \le \frac{\epsilon}{4} .
    $$
    Simliarly, by assumption we can choose a $\delta$ for which
    $$
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
        \le \frac{\epsilon}{C} .
    $$
    If we choose $C$ and $\delta$ that do this for both $X^{(n)}$ and $Y^{(n)}$,
    then each of the terms in equation~\eqref{eqn:xy_bound} are bounded by $\epsilon/4$,
    and condition (2) is satisfied for the product process.
\end{proof}

\begin{lemma}
    \label{lem:test_fn_tightness}
    Let $f$ be a bounded, continuous real-valued function on $\IR^d \times [0, \infty)$
    with uniformly bounded first and second derivatives
    for which there exists a $u_0$ such that if $u > u_0$ then $f(x, u) = 0$.
    Then, the sequence of real-valued processes $(\langle f, \lp_t^N \rangle)_{t \ge 0}$
    are tight in the space of c\`adl\`ag processes.
\end{lemma}

    It is evident from the description of the process
    (or, by differentiating in Definition~\ref{defn:lookdown_mgale})
    that
    \begin{align*}
    \begin{split}
    M^f_t
        &=
        \langle f, \lp_t^N \rangle
        -
        \langle f, \lp_0^N \rangle
    \\ & \qquad \qquad {}
        -
        \int_0^t
        \big\langle
            a^N(x, \eta^N_s)
            \int_u^N \int_{\IR^d}
            \left(
                f(y, u_1) + f(x, u_1) + f(y, u) - f(x, u)
            \right)
            du_1 q^m(x, dy, \eta^N_s)
    \\ & \qquad \qquad \qquad \qquad {}
            +
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            \frac{d}{du} f(x, u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
    \end{split}
    \end{align*}
    is a martingale,
    with quadratic variation
    \begin{align*}
        [M^f]_t
        &=
        \int_0^t
        \langle a^N(x, \eta_s^N) \int_u^N \int_{\IR^d}
            f(y, u_1)^2 +
            \left( f(x, u_1) + f(y, u) - f(x, u) \right)^2
        du_1 q^m(x, dy, \eta^N_s)
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}

    Now we would like to bound the terms in the integral,
    to arrive at a bound using Gronwall's inequality, as before.

    First, note that there exists a constant $C_a$
    such that $a^N(x, \eta) \le C_a$
    (since $\gamma$ and $r$ are bounded above, and $\theta/N \to \alpha \ge 0$),
    and for any $\epsilon > 0$ and $T$ there is a $C_b$ such that for all $N$,
    $$
        \IP\left\{
            \sup_{0 \le s \le T} \sup_x |b^N(x, \eta_s^N)| > C_b
        \right\}
        < \epsilon 
    $$
    \comment{reference where we show this above}.
    Now, since $f$ vanishes for $u \ge u_0$,
    on the event that $\sup_x |b^N(x, \eta_s^N)| \le C_b$ then
    if $N \ge u_0$,
    \begin{align*}
    &
        \int_{\IR^d} \int_0^N
            \left(
                a^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            \frac{d}{du} f(x, u)
        \lp_s^N(dx, du)
    \\  & \qquad
    \le
        \int_{\IR^d} \int_0^{u_0}
            \left(
                C_a u_0^2
                + C_b u_0
            \right)
            \left\| \frac{d}{du} f \right\|_\infty
        \lp_s^N(dx, du) .
    \\  & \qquad
    \le
        C_0 \lp_s^N(\IR^d \times [0, u_0]) .
    \end{align*}

    Putting this together,
    \begin{align*}
        \IE\left[
            \langle f, \lp_t^N \rangle
        \right]
        \le
    \end{align*}

\subsubsection{OLD STUFF REMOVE}

@@
